Union and intersection contracts are hard, actually

### 2021-09-14 paper.tex
Cut some text on effects, purity and CSE

Higher-order contracts described in Section~\ref{sub-sec:contracts} already
introduce side-effects: semantically, the \nickel{blame} operator signaling
contract violation is equivalent to non-termination (error reporting aside).
But non-termination is a special kind of effect, if only because it preserves
referential transparency. In a lazy language like Nickel,
(\ref{eq:cbn-expansion}) still holds for all terms in the presence of
non-termination. In a strict language (call-by-value), (\ref{eq:cbn-expansion})
fails in all generality because the transformation may cause a previously
unevaluated term to be suddenly evaluated.  However, (\ref{eq:cbn-expansion})
holds given slight restrictions on the context $M$: for example, it holds as
long as at least one occurrence of $x$ in $M$ is in scope (i.e. not inside a
function). Beside, many languages out there (including Nickel) already feature
non-termination. For such languages, adding contracts does not change what
optimizations are valid.

Strikingly, we will see in Section~\ref{sec:issues-sem} that the introduction of
union and intersection contracts \emph{breaks referential transparency}. Unions and
intersections make (\ref{eq:cbn-expansion}) unsound even in a lazy language,
preventing the kind of optimization of
Figure~\ref{fig:optimizations-let-floating-ex} to fire in general. In a strict
language, the restrictions to (\ref{eq:cbn-expansion}) outlined above are not
sufficient anymore. If the effect of plain higher-order contracts behaves as
non-termination, the effect of union and intersection contracts rather behaves
as shared state.

% TODO Will this be demostrated?
% yes, I think is better untold, and the paragraph starts by
% referencing section 5

>>>>

Strikingly, we will see in Section~\ref{sec:issues-sem} that the introduction of
union and intersection contracts \emph{breaks referential transparency} and make
(\ref{eq:cbn-expansion}) \emph{unsound}\footnote{(\ref{eq:cbn-expansion}) holds
  even in a technically impure language like Nickel with non-termination and
  higher-order contracts because of lazy evaluation. This is not the case in a strict language,
  although in that case, (\ref{eq:cbn-expansion}) still holds more often without
  unions and intersections than with.}, preventing the kind of optimization of
Figure~\ref{fig:optimizations-let-floating-ex} to fire in general.
